True versus apparent shapes of astrophysical bow shocks

### 2016-06-30 papers/Paper1/sec-conic.tex
Corrected/updated some equations.
+ Corrected bad written equations in sections 4,5
+ Updated equation for radius of curvature
Found and corrected more typos in quadrics section
Improve Fig 7 (family of conics)

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure}
\setlength\tabcolsep{0pt}
\begin{tabular}{cc}
\includegraphics[height=0.54\linewidth, trim=30 0 30 0, clip]{ellipse_py_lab} &
\includegraphics[height=0.54\linewidth, trim=60 0 70 0, clip]{hyperbola_py_lab}
\end{tabular}
\label{fig:conics}
\caption{Schematic representation of conic like bowshocks: A) Ellipse and B) Hyperbola The center of the conic is located at a distance $x_0=R_0-a$ from the origin at the $x$ axis. 
The radius of curvature at the axis is $R_c$, and $R_{90}$ is t he radius of the conic in the perpendicular direction to $R_0$.
}
\end{figure}
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{A whole set of conics with the same radius of curvature. The parameter $\theta_c$ is used as the equivalent of the excentricity. This angle is defined as positive for closed curves
and negative for the opened ones. An special case is $\theta_c=0$, which describes a parabola.}
\label{fig:conics-family}
\end{figure}

Due to the similarities between the parametrization of the ellipse and the hyperbola, we can do the following parametrization:
\begin{align}
x = a\Cos(t)-x_0 \label{eq:par-x}\\ 
y = b\Sin(t) \label{eq:par-y}
\end{align}
where:
\begin{align}
\Cos(t) = \left\lbrace \begin{array}{c}
\cos t ~\mathrm{if~ellipse} \\
\cosh t ~\mathrm{if~hyperbola}
\end{array}\right.\\
\Sin(t) = \left\lbrace \begin{array}{c}
\sin t ~\mathrm{if~ellipse} \\
\sinh t ~\mathrm{if~hyperbola}
\end{array}\right. \\
-\pi < t < \pi \\
R_0 = a - x_0 \\
a<0 ~ \mathrm{for~hyperbola}
\end{align}
We have set a whole family of curves with a singular parameter denoted as $\tan\theta_c \equiv \frac{b}{a}$. Positive values of this parameter describe closed curves
(i.e. ellipses) and the negative ones describe open curves (i.e. hyperbolas). Particular cases are $\theta_c =\frac{\pi}{4}$, which describes a circle and $\theta_c=0$, which describes
a parabola.
To determine uniquely a given conic in terms of measurable quantities, we use the set of characteristic radii $(R_0,R_c,R_{90})$ in such way the conic parameters $(a,b,\theta_c)$
could be derived from the former. The transformation between the conic parameters and the characteristic radii are given by:
\begin{align}
\frac{R_c}{D} &= \frac{b^2}{a} \label{eq:Rc-conic}\\ 
\frac{R_{90}}{D} &= b\left(1 \mp \frac{(R_0-a)^2}{a^2}\right)
\end{align}
The inverse transformation is given by:
\begin{align}
a &= \frac{R_c}{2A-B^2} \label{eq:a-conic}\\
b &= \left(aR_c\right)^{1/2}\\
\theta_c &= \left|B^2 - 2A\right|^{1/2} \label{eq:thc-conic} \\
\mathrm{where:}& B\equiv \frac{R_{90}}{R_0} \\
A \equiv \frac{R_c}{R_0}
\end{align}
From equations (\ref{eq:a-conic}) we can see that for $R^2_{90} > 2 R_c$, the conic parameter $a$ is negative, so in this case the resulting conic is an hyperbola, and we must take the negative root
in equation (\ref{eq:thc-conic}). The opposite applies for ellipses.
\subsection{Plane-of-sky projection of quadric surfaces} 
\begin{figure*}
\includegraphics[width=0.9\linewidth]{annotated}
\caption{Projected Radius of curvature vs projected $R_0$ for $\theta_c=60^\circ,45^\circ,28^\circ, 0^\circ,-15^\circ,-30^\circ$. Each colored points represent
the same shape at a different inclination. Being the dark blue the ones with the lower inclination and the red ones with the highest value. For closed shapes, there is
an assymptotic limit for the projected radius of curvature at high inclinations, while for open shapes there is an upper limit for inclination where a tangent line can be observed.}
\label{fig:Apqp}
\end{figure*}

@@

We could work in another frame where the origin is located at the conic's center: $(X,Y)$, where $X=x-X_0$ and $Y=y$.
In this frame,  we use the transformations (\ref{eq:Trans})  to obtain the apparent shape of the shell:
\begin{align}
X' = \frac{\Cos(t)}{a\cos i}\left(a^2\cos^2 i \pm b^2\sin^2 i\right)  \label{eq:conic-projected-x}\\
Y'= b\Sin(t)\left(1-\frac{b^2}{a^2}\tan^2 i\Cot(t)\right)^{1/2}
\label{eq:conic-projected-y}
\end{align}

@@

Comparing, (\ref{eq:conic-projected-x}), (\ref{eq:conic-projected-x-2}), (\ref{eq:conic-projected-y}) and (\ref{eq:conic-projected-y}) we find that
\begin{align}
a' = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} \\
\Cos(t') = \frac{\Cos(t)}{a\cos i} \\
b' = b
\end{align} 
Which proves that effectively the projection of a given conic is the same conic. With this knowledge, we can obtain the characteristic radii in
the observer's frame.
\begin{align}
D' = D\cos i \\
R'_0 = a' - (a-R_0)\cos i \\
R'_0 = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2}  - (a-R_0)\cos i
\end{align}

With  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ and introducing the
radius of curvature $R_c\equiv \frac{b^2}{a}$ we find that:
\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1+A\cot^2\theta_c(f(i;\theta_c)-1) \right)
\end{align}
Where $A\equiv \frac{R_c}{R_0}$

>>>>

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure}
\setlength\tabcolsep{0pt}
\begin{tabular}{cc}
\includegraphics[height=0.54\linewidth, trim=30 0 30 0, clip]{ellipse_py_lab} &
\includegraphics[height=0.54\linewidth, trim=60 0 70 0, clip]{hyperbola_py_lab}
\end{tabular}
\label{fig:conics}
\caption{Schematic representation of conic like bowshocks: A) Ellipse and B) Hyperbola The center of the conic is located at a distance $x_0=R_0-a$ from the origin at the $x$ axis. 
The radius of curvature at the axis is $R_c$, and $R_{90}$ is t he radius of the conic in the perpendicular direction to $R_0$.
}
\end{figure}
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{Example of a family of conic sections, all with the same
  radius of curvature at their apex (marked by white dot) on the
  \(x\)~axis: \(R_c = 2 R_0\). The parameter
  $\theta_c = \pm \tan^{-1} b/a$ describes the family, with negative
  values corresponding to hyperbolae.}
\label{fig:conics-family}
\end{figure}

Due to the similarities between the parametrization of the ellipse and the hyperbola, we can do the following parametrization:
\begin{align}
x = a\Cos(t)-x_0 \label{eq:par-x}\\ 
y = b\Sin(t) \label{eq:par-y}
\end{align}
where:
\begin{align}
\Cos(t) = \left\lbrace \begin{array}{c}
\cos t ~\mathrm{if~ellipse} \\
\cosh t ~\mathrm{if~hyperbola}
\end{array}\right.\\
\Sin(t) = \left\lbrace \begin{array}{c}
\sin t ~\mathrm{if~ellipse} \\
\sinh t ~\mathrm{if~hyperbola}
\end{array}\right. \\
-\pi < t < \pi \\
R_0 = a - x_0 \\
a<0 ~ \mathrm{for~hyperbola}
\end{align}
We have set a whole family of curves with a singular parameter denoted as $\tan\theta_c \equiv \frac{b}{a}$. Positive values of this parameter describe closed curves
(i.e. ellipses) and the negative ones describe open curves (i.e. hyperbolas). Particular cases are $\theta_c =\frac{\pi}{4}$, which describes a circle and $\theta_c=0$, which describes
a parabola.
To determine uniquely a given conic in terms of measurable quantities, we use the set of characteristic radii $(R_0,R_c,R_{90})$ in such way the conic parameters $(a,b,\theta_c)$
could be derived from the former. The transformation between the conic parameters and the characteristic radii are given by:
\begin{align}
\frac{R_c}{D} &= \frac{b^2}{a} \label{eq:Rc-conic}\\ 
\frac{R_{90}}{D} &= b\left[\pm\left(1 - \frac{(R_0-a)^2}{a^2}\right)\right]^{1/2}
\end{align}
The inverse transformation is given by:
\begin{align}
\frac{a}{R_0} &= \frac{A}{2A-B^2} \label{eq:a-conic}\\
\frac{|b|}{R_0} &= \frac{A}{\left|2A-B^2\right|}\\
\tan\theta_c &= \left|B^2 - 2A\right|^{1/2} \label{eq:thc-conic} \\
\mathrm{where:}& B\equiv \frac{R_{90}}{R_0} \\
A \equiv \frac{R_c}{R_0}
\end{align}
From equations (\ref{eq:a-conic}) we can see that for $R^2_{90} > 2 R_c$, the conic parameter $a$ is negative, so in this case the resulting conic is an hyperbola, and we must take the negative root
in equation (\ref{eq:thc-conic}). The opposite applies for ellipses.
\subsection{Plane-of-sky projection of quadric surfaces} 
\begin{figure*}
\includegraphics[width=0.9\linewidth]{annotated}
\caption{Projected Radius of curvature vs projected $R_0$ for $\theta_c=60^\circ,45^\circ,28^\circ, 0^\circ,-15^\circ,-30^\circ$. Each colored points represent
the same shape at a different inclination. Being the dark blue the ones with the lower inclination and the red ones with the highest value. For closed shapes, there is
an assymptotic limit for the projected radius of curvature at high inclinations, while for open shapes there is an upper limit for inclination where a tangent line can be observed.}
\label{fig:Apqp}
\end{figure*}

@@

We could work in another frame where the origin is located at the conic's center: $(X,Y)$, where $X=x-X_0$ and $Y=y$.
In this frame,  we use the transformations (\ref{eq:Trans})  to obtain the apparent shape of the shell:
\begin{align}
X' = \frac{\Cos(t)}{a\cos i}\left(a^2\cos^2 i \pm b^2\sin^2 i\right)  \label{eq:conic-projected-x}\\
Y'= b\Sin(t)\left(1-\frac{b^2}{a^2}\tan^2 i\Cot^2(t)\right)^{1/2}
\label{eq:conic-projected-y}
\end{align}

@@

Comparing, (\ref{eq:conic-projected-x}), (\ref{eq:conic-projected-x-2}), (\ref{eq:conic-projected-y}) and (\ref{eq:conic-projected-y}) we find that
\begin{align}
a' = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} \\
\Cos(t') = \frac{a'\Cos(t)}{a\cos i} \\
b' = b
\end{align} 
Which proves that effectively the projection of a given conic is the same conic. With this knowledge, we can obtain the characteristic radii in
the observer's frame.
\begin{align}
D' = D\cos i \\
R'_0 = a' - (a-R_0)\cos i \\
R'_0 = \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2}  - (a-R_0)\cos i
\end{align}

With  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ we find that:
\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1+A\cot^2\theta_c(f(i;\theta_c)-1) \right)
\end{align}
Where $A\equiv \frac{R_c}{R_0}$
